For submission to the Journal of Integrated Pest Management

### 2021-06-07 editorial_reviews/response_editors_ofv_fife_2021.Rmd

Table 1:
L338: “…with symptoms of orchid fleck virus infection found in…”
I am a little concerned about reporting on plants with symptoms, vs confirmed by diagnostics.  As mentioned earlier, different dichorhaviruses can cause similar symptoms in the same host.  Although remote, there is the possibility another virus could be responsible for the symptoms in a different location

Figure captions:
Please use “orchid fleck virus (OFV)” when describing a physical virus, and "Orchid fleck dichorhavirus" (in italics) when describing the species.

**Response to Reviewer 1:**
**Thank you for reviewing our article. We have made the following changes based on your remarks:**

**We have adjusted some wording, organization, and grammar of the manuscript in an attempt to aid clarity and accuracy. We also reviewed and corrected some citations.**

**We rewrote the sections describing sampling methods and diagnostics in the abstract and main text. Both samples from Leon County and Alachua were found to be positive for OFV; only samples from Leon were examined for mites. We have removed the claim that the virus is widespread.**

Specific Edits:

- Added discussion on why OFV has not been reported in nearby citrus.
- We have removed all references to cileviruses
- We have corrected inconsistencies with taxonomic authorities
- We have not added a discussion of potential control measures due to word limits, and lack of specific documentation regarding control methods/chemical applications for these specific ornamental plants.

Also: 
L33: “Orchid fleck virus (OFV) is a dichorhavirus that infects over 50 plant species belonging to the families Asparagaceae, Orchidaceae, and Rutaceae.  The only known vectors for…” $\rightarrow$ **rewritten**

L37: (Liriope and Ophiopogon spp.) $\rightarrow$ **fixed**

L40: “…Sanger sequencing.” $\rightarrow$ **fixed**

L41: remove “the known” and “the NCBI” $\rightarrow$ **fixed**

L42: I see no mention of a qPCR assay in the main text $\rightarrow$ **fixed**

L43: “…possibly infected plants of the family Asparagaceae from…” $\rightarrow$ **rewritten**

L44: please elaborate on “Identification of partial genome sequence…” is this sequence from the PCR products or was HTS performed on the samples? $\rightarrow$ **rewritten**

L48: perhaps identify what s.l. means; authorities are provided elsewhere in the manuscript $\rightarrow$ **sensu lato in this case refers to the fact that it not the specific mite referred to by Baker as *B. californicus*, which has been split into 3 subgroups, but have not been adequately described yet**

L49: italicize “Brevipalpus” $\rightarrow$ **fixed**

L59: Virus species should not be abbreviated; remove “(OFV)” $\rightarrow$ **Is this an acceptable alternative?**

L60: virus families are italicized (Rhabdoviridae) $\rightarrow$ **fixed**

L63: replace “sole” with “only known” $\rightarrow$ **fixed**

L64: “group of mites” vs “sensu lato”? $\rightarrow$ **fixed**

L69: I don’t think referencing figures 1 and 2 are appropriate here $\rightarrow$ **fixed**

L73,76: is listing the second author the correct way to distinguish articles for this journal?  Rather than “a” or “b” after publication year? $\rightarrow$ **citation manager error, fixed**

L77: remove “: (Citrus) $\rightarrow$ **fixed**

L78: Olmedo-Velarde et al. is not an appropriate reference here $\rightarrow$ **citation should have been @Velarde2021, fixed**

L82: “In June 2020,…”; (Lirope spp.) $\rightarrow$ **fixed**

L83: belongs $\rightarrow$ **fixed**

L85: perhaps “monocotyledonous liliod plants native to southeastern Asia (Chase…” $\rightarrow$ **fixed**

L86: is it appropriate to provide an authority for the genus (Ker Gawler)? $\rightarrow$ **fixed**

L92: none of these virus names should be italicized; INSV and TSWV are orthotospoviruses – why mentioned separately? $\rightarrow$ **I believe that these were specific tests, fixed**

L97: “…as well as Aspidistra elaitor Blume (Asparagaceae), which was suspected…” $\rightarrow$ **fixed, *A. elatior***

L103: “Lirope sp.” $\rightarrow$ **fixed**

L108: how many sites? $\rightarrow$ **There are multiple sites with OFV-infected plants in Leon County, and one site identified in Alachua County. The plant samples tested from Leon Co. were taken from the first site where the pathogen was found.**

L127: This is incorrect: Blanchfield et al 2001 identified a virion larger than OFV, and the sample was negative for OFV when tested by RT-PCR. $\rightarrow$ **Thank you for correcting this mistake, fixed**

L134: Nolinoideae $\rightarrow$ **fixed**

L145: The ICTV is not responsible for classification below species level, and as such, does not recognize virus strains; scientists and experts can recognize strains, but the  ICTV does not. $\rightarrow$ **fixed**

L152: “including Vanilla spp. in southern Florida…” $\rightarrow$ **clarified**

L154: a reference for this statement is advisable  $\rightarrow$ **clarified**

L155: “…found that the virus which previously affected…”  $\rightarrow$ **clarified**

L165: “…vectors of dichorhaviruses (Knorr 1968,…” $\rightarrow$ **clarified**

L168: “…identify the vector(s) of OFV in Florida…” $\rightarrow$ **clarified**

Table 1:
L338: “…with symptoms of orchid fleck virus infection found in…”
I am a little concerned about reporting on plants with symptoms, vs confirmed by diagnostics.  As mentioned earlier, different dichorhaviruses can cause similar symptoms in the same host.  Although remote, there is the possibility another virus could be responsible for the symptoms in a different location $\rightarrow$ **clarified in text**

Figure captions:
Please use “orchid fleck virus (OFV)” when describing a physical virus, and "Orchid fleck dichorhavirus" (in italics) when describing the species. $\rightarrow$ **fixed**

#### Reviewer: 2

@@

My primary suggestion is that they be more specific about referring that the disease was reported in Florida. Other regions of the country were not part of the current work and should not be suggested as having a broad distribution.

I made a few editions, suggesting proper spellings for species “Family” names and the use of correct reference directly on the manuscript file attached.

Kitajima et. al. (2011) - correct reference mentioned is:
Kitajima E W, Chagas CM, Harakava R, Calegario RF, Freitas-Astúa J, Rodrigues JCV, Childers, CC. 2011. Citrus Leprosis in Florida, USA, appears to have been caused by the Nuclear Type of Citrus Leprosis Virus (CiLV-N). Virus Reviews & Research 16:1-5.

**Reply to Reviewer 2:**
**Thank you for reviewing our article. We have made the following changes based on your remarks:**

**We have removed references to a 'wide distribution' in Florida, corrected species names and inconsistencies with citations. We are grateful for your assistance.**

Specific Edits:

- fixed reference manager errors: added brackets {} around capitalized words {} and italic species names '{\\textit{}}' to protect case. Also: Asparagalean, Amaryllidaceae, Asparagaceae, Xanthorrhoeaceae, *Rhabdoviridae*, Asparagaceae, RNA, *Psilanthus*

Italicized all instances of *Rhabdoviridae*

- L321 - We do not believe it is necessary to italicize the plant family 'Asparagaceae'

### 2021-06-07 ofv_fife_2021.Rmd
Made final edits to MS, ready to forward to rest of the authors

Chlorotic ringspots and flecking was seen affecting Liriopogons (*Liriope* spp. and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) and USDA-ARS for identification.

@@

### Keywords:
False spider mite, flat mite, *Brevipalpus*-transmitted viruses, *Liriope*, Nolinoideae, *Ophiopogon*, *Aspidistra*, Ruscaceae, Rutaceae, Asparagaceae, orchid, Orchidaceae, pests, ornamental plants, Orchid fleck virus.

\pagebreak
*Orchid fleck dichorhavirus* (OFV) is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae) are the only known vectors of dichorhaviruses [@Maeda1998]. *Brevipalpus californicus* (Banks) sensu lato mites are the only group known to transmit OFV in a persistent propagative manner [@Kondo2003].

@@

OFV naturally infects more than fifty species of Orchidaceae [@Kitajima2010; @Peng2013], some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], and Rutaceae: (*Citrus*), where it causes citrus leprosis-like symptoms [@Cook2019; @OlmedoVelarde2019; @Roy2015; @Roy2020]. Mechanical transmission of OFV is possible under laboratory conditions to some plants belonging to the plant families Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

During June 2020, chlorotic ringspot symptoms were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, comprised of grass-like monocotyledonous liliod plants native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* Ker Gawler (Asparagaceae: Nolinoidaea) are considered the most important ground cover plant in the southeastern United States [@Mcharo2003].

Viral infections of suspected leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples were tested with one step conventional RT-PCR, and were found negative for begomovirus, carlavirus, potyvirus, tospovirus, and *Cucumber mosaic virus*, *Impatiens necrotic spot virus*, *Tobacco mosaic virus*, and *Tomato spotted wilt virus*.

@@

The flat mites collected were initially suspected to belong to *B. californicus* s. l. after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s. l. species group, sensu @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. With this in mind, new mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Baker and *B. confusus* Donnadieu (`r citef("fig_4")`).

The first report of OFV in the US is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. In 2001, @Blanchfield2001 published the first certain report of OFV from Hawaii. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015].

@@

OFV consists of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in gene order, content, and genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. The *Brevipalpus* mites collected from liriopogons and *A. elatior* in Leon county were abundant on OFV-infected plants very near to citrus trees, some plants even surrounding the trunk. *B. californicus* s. l. have been reported as a pest of citrus [@Childers2003] and are often collected from citrus rinds [@Baker1949; @Baker1987], so the proximity of these mite vectors to citrus raises the question: why these trees are not currently infected with OFV-Orc? Each possibility suggest further research, but there are a few possible explanations. It is important to note thge uncertaintly surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. californicus* (the most likely culprit), *B. obovatus*, and *B. confusus*. Each species has its own unique biology and all have been implicated with a variety of different hosts. This suggests that the spread of OFV-Orc would be a function of various combinations of a number of potential factors, possibly including: host preferences, vectorial capacity, viral propagation/circulation in the vector, viral acquisition times, and feeding times required for transmission. Another possiblity is that the *B. californicus* which we find on liriopogons and *A. elatior* are not actually the same species as those found on citrus, and instead represent 2-3 different cryptic species.

Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the US multi-billion dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of Dichorhaviruses (OFV) [@Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`), and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to determine which of these species are responsible for transmission of OFV-Orc, and therefore which mite populations need to be monitored or controlled. By doing so, we can determine the risk they represent for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

### Figure captions
`r fig_nums(name = "fig_1", caption = "Variety of symptoms expressed by *Liriope* spp. infected with Orchid fleck dichorhavirus (OFV): (a) ringspot symptoms on *Liriope gigantea* (b-c) Details of ringspot symptoms on *Liriope gigantea* (d) chlorotic ringspot *Liriope muscari* cv. 'Silvery Sunproof'")`

`r fig_nums(name = "fig_2", caption = "Symptoms expressed by *Aspidistra elatior* infected with Orchid fleck dichorhavirus (OFV): (a) Detail of leaf chlorosis (b) Chlorosis caused by OFV appears similar to sunburn damage (c-d) Chlorotic ringspot may indicate early symptoms of OFV")`

>>>>

Chlorotic ringspots and flecking was seen affecting Liriopogons (*Liriope* and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) and USDA-ARS for identification.

@@

### Keywords:
False spider mite, flat mite, *Brevipalpus*-transmitted viruses, *Liriope*, Nolinoidaea, *Ophiopogon*, *Aspidistra*, Ruscaceae, Rutaceae, Asparagaceae, orchid, Orchidaceae, pests, ornamental plants, Orchid fleck virus.

\pagebreak
*Orchid fleck dichorhavirus*, commonly referred to as Orchid Fleck Virus (OFV), is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*. OFV is a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae) are the only group known to transmit dichorhaviruses [@Maeda1998], and *Brevipalpus californicus* (Banks) sensu lato are the only mites which do so in a persistent propagative manner [@Kondo2003].

@@

OFV naturally infects more than fifty species of Orchidaceae [@Kitajima2010; @Peng2013], some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], and Rutaceae, where it causes citrus leprosis-like symptoms [@Cook2019; @Velarde2021; @Roy2015; @Roy2020]. Mechanical transmission of OFV is possible under laboratory conditions to some plants belonging to the plant families Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

During June 2020, chlorotic ringspot symptoms were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, comprised of grass-like monocotyledonous liliod plants native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* (Asparagaceae: Nolinoidaea) are considered the most important ground cover plant in the southeastern United States [@Mcharo2003].

Viral infections of suspected leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples were tested with one step conventional RT-PCR, and were found negative for begomovirus, carlavirus, potyvirus, tospovirus, Cucumber mosaic virus and Tobacco mosaic virus.

@@

The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, sensu @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. With this in mind, new mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Baker and *B. confusus* Donnadieu (`r citef("fig_4")`).

The first report of OFV in the US is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015].

@@

OFV consists of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in gene order, content, and genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. The *Brevipalpus* mites collected from liriopogons and *A. elatior* in Leon county were abundant on OFV-infected plants very near to citrus trees, some plants even surrounding the trunk. *B. californicus* s. l. have been reported as a pest of citrus [@Childers2003] and are often collected from citrus rinds [@Baker1949; @Baker1987], so the proximity of these mite vectors to citrus raises the question: why these trees are not currently infected with OFV-Orc? Each possibility suggest further research, but there are a few possible explanations. It is important to note the uncertainty surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. californicus* (the most likely culprit), *B. obovatus*, and *B. confusus*. Each species has its own unique biology and all have been implicated with a variety of different hosts. This suggests that the spread of OFV-Orc would be a function of various combinations of a number of potential factors, possibly including: host preferences, vectorial capacity, viral propagation/circulation in the vector, viral acquisition times, and feeding times required for transmission. Another possiblity is that the *B. californicus* which we find on liriopogons and *A. elatior* are not actually the same species as those found on citrus, and instead represent 2-3 different cryptic species.

Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern strains of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the US multi-billion dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of dichorhaviruses (OFV) [@Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`), and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to determine which of these species are responsible for transmission of OFV-Orc, and therefore which mite populations need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

### Figure captions
`r fig_nums(name = "fig_1", caption = "Variety of symptoms expressed by *Liriope* spp. infected with *Orchid fleck dichorhavirus*: (a) ringspot symptoms on *Liriope gigantea* (b-c) Details of ringspot symptoms on *Liriope gigantea* (d) chlorotic ringspot *Liriope muscari* cv. 'Silvery Sunproof'")`

`r fig_nums(name = "fig_2", caption = "Symptoms expressed by *Aspidistra elatior* infected with *Orchid fleck dichorhavirus*: (a) Detail of leaf chlorosis (b) Chlorosis appears similar to sun damage (c-d) Chlorotic ringspot may indicate early symptoms of OFV")`
